---
title: "Hands-on_Ex02"
---
2.3.1 & 2.3.2: Data & Importing of Data into R
```{r}
pacman::p_load(sf, tmap, tidyverse)
```

```{r}
data_folder_mp <- "../../data/MasterPlan2014"

mpsz <- st_read(dsn = data_folder_mp,
                 layer = "MP14_SUBZONE_WEB_PL")

mpsz
```
2.3.3 Importing Attribute Data into R
```{r}
data_folder_rp <- "../../data/Respopagesextod2011to2020"
data_file_rp <- paste0(data_folder_rp, "/respopagesextod2011to2020.csv")
popdata <- read_csv(data_file_rp)

```
2.3.4 Data Preparation

2.3.4.1 Data wrangling
```{r}
popdata2020 <- popdata %>%
  filter(Time == 2020) %>%
  group_by(PA, SZ, AG) %>%
  summarise(`POP` = sum(`Pop`)) %>%
  ungroup()%>%
  pivot_wider(names_from=AG, 
              values_from=POP) %>%
  mutate(YOUNG = rowSums(.[3:6])
         +rowSums(.[12])) %>%
mutate(`ECONOMY ACTIVE` = rowSums(.[7:11])+
rowSums(.[13:15]))%>%
mutate(`AGED`=rowSums(.[16:21])) %>%
mutate(`TOTAL`=rowSums(.[3:21])) %>%  
mutate(`DEPENDENCY` = (`YOUNG` + `AGED`)
/`ECONOMY ACTIVE`) %>%
  select(`PA`, `SZ`, `YOUNG`, 
       `ECONOMY ACTIVE`, `AGED`, 
       `TOTAL`, `DEPENDENCY`)
```
2.3.4.2 Joining the attribute data and geospatial data
```{r}
popdata2020 <- popdata2020 %>%
  mutate_at(.vars = vars(PA, SZ), 
          .funs = list(toupper)) %>%
  filter(`ECONOMY ACTIVE` > 0)

mpsz_pop2020 <- left_join(mpsz, popdata2020,
                          by = c("SUBZONE_N" = "SZ"))
dir.create("../../data/rds", showWarnings = FALSE)
write_rds(mpsz_pop2020, "../../data/rds/mpszpop2020.rds")
```
2.4 Choropleth Mapping Geospatial Data Using tmap

2.4.1 Plotting a choropleth map quickly by using qtm()
```{r}
tmap_mode("plot")
qtm(mpsz_pop2020, 
    fill = "DEPENDENCY")
```
2.4.2 Creating a choropleth map by using tmap’s elements
```{r}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY", 
          style = "quantile", 
          palette = "Blues",
          title = "Dependency ratio") +
  tm_layout(main.title = "Distribution of Dependency Ratio by planning subzone",
            main.title.position = "center",
            main.title.size = 1.2,
            legend.height = 0.45, 
            legend.width = 0.35,
            frame = TRUE) +
  tm_borders(alpha = 0.5) +
  tm_compass(type="8star", size = 2) +
  tm_scale_bar() +
  tm_grid(alpha =0.2) +
  tm_credits("Source: Planning Sub-zone boundary from Urban Redevelopment Authorithy (URA)\n and Population data from Department of Statistics DOS", 
             position = c("left", "bottom"))
```
2.4.2.1 Drawing a base map
```{r}
tm_shape(mpsz_pop2020) +
  tm_polygons()
```
2.4.2.2 Drawing a choropleth map using tm_polygons()
```{r}
tm_shape(mpsz_pop2020)+
  tm_polygons("DEPENDENCY")
```
2.4.2.3 Drawing a choropleth map using tm_fill() and *tm_border()**
```{r}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY")
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY") +
  tm_borders(lwd = 0.1,  alpha = 1)
```
2.
```{r}

```
2.
```{r}

```
2.
```{r}

```
2.
```{r}

```
2.
```{r}

```
2.
```{r}

```
2.
```{r}

```
2.
```{r}

```
2.
```{r}

```
2.
```{r}

```
2.
```{r}

```
2.
```{r}

```
2.
```{r}

```
2.
```{r}

```
2.
```{r}

```
2.
```{r}

```
